Thesis outline

### 2017-06-05 KellyDouglass/Thesis/Thesis_sections/smallScaleEnvironment.tex
GV chapter
Small-scale theory, data

@@

@@

@@

@@

\subsection{Absolute distance to nearest neighbor}

@@

% SDSS

@@

@@

\cite{Berlind06}

\subsection{Finding the nearest neighbor}
% Absolute magnitude distribution of neighbors

@@

\subsection{Selection effects}

@@

@@

\section[Conclusion]{Conclusion}

@@

>>>>

Morphology-density relation \cite{Dressler80}

@@

We employ two different spatial metrics to determine the best way to locate a 
galaxy's nearest neighbor.  Within a relative velocity 
$v_{rel} < 300\text{ km/s}$, we use the physically closest (in units of \hMpc) 
and the neighbor with the smallest fraction of its virial radius; both these 
methods are described below in further detail.  The two methods result in 
different neighboring galaxies for 40\% of our dwarf galaxy sample.

@@

A galaxy's redshift is composed of both the expansion of the universe and the 
galaxy's peculiar velocity (its motion relative to its surrounding galaxies and 
environment).  Known as the ``Finger of God'' effect, the peculiar velocity 
causes galaxy groups and clusters to appear extended along the line of sight 
when using redshift as a distance measurement.  When we calculate distances on 
large scales, the peculiar velocity is negligible.  However, when we study the 
distance between two objects in the universe on a small scale (within a few 
\hMpc), the peculiar velocity dominates.  The relative velocity $v_{rel}$ 
between galaxies $a$ and $b$ is defined as
\begin{equation}
    v_{rel} = |z_a - z_b|c
\end{equation}
where $z$ is the redshift.  We require all 
neighbors to have a maximum relative velocity $v_{rel} < 300\text{ km/s}$.  We 
show in Sec. \ref{sec:selection_effects} that our results are insensitive to 
this value.

@@

The projected distance between two galaxies is found by projecting the two 
galaxies onto a sphere of radius $\bar{r}$, where $\bar{r}$ is the average 
distance from Earth of the two galaxies.  This assumes that two galaxies which 
are gravitationally bound are in a system where the relative velocity between us 
and it is 0 and that $\bar{r}$ is the distance to the center of the system.  The 
distance to the system is calculated as
\begin{equation}
    \bar{r} = \frac{\bar{z}c}{H_0}
\end{equation}
where $\bar{z}$ is the average redshift of the two galaxies, $c$ is the speed of 
light, and $H_0$ is the Hubble constant.  Each galaxy's right ascension (RA) and 
declination (dec.) are converted to Cartesian coordinates to find their sky 
separation.

@@

From \cite{Hwang10}, we calculate the virial radius of a galaxy to be
\begin{equation}\label{eqn:r_vir}
    r_{vir} = (3L\gamma/4\pi/200\rho_c)^{1/3}
\end{equation}
where $L$ is the galaxy's luminosity, $\gamma$ is the mass-to-light ratio, and 
$\rho_c$ is the critical density of the universe.  The critical density is the 
density at which evenly distributed gas would collapse to form a galaxy.  The 
mass-to-light ratio depends on the galaxy's morphology type, where 
$\gamma_{early} = 2\gamma_{late}$.  \cite{Hwang10} defines 
$200\rho_c = 740\bar{\rho}$, where 
$\bar{\rho} = (0.0223\pm 0.0005)(\gamma_{late}L_{-20})(\text{\hMpc})^3$.  We can 
then rewrite Eqn. \ref{eqn:r_vir} as
\begin{equation}
    r_{vir} = \left( \frac{3}{4\times 740\times 0.0223\pi} \frac{\gamma}{\gamma_{late}} \frac{L}{L_{-20}} \right)^{1/3}
\end{equation}
where $L_{-20}$ is the luminosity of a galaxy with $M_r = -20$.  We can then use 
the virial radius to scale the distance between two galaxies as a fraction of 
the neighbor's virial radius.

Rather than using a measure of the virial radius for the group analysis, we 
scale the distance as a fraction of the group's rms radius.

\subsection{Absolute distance to nearest neighbor}
% Including redshift

We can also calculate the distance between two galaxies by ignoring the effects 
of their relative velocities.

@@

% SDSS
The Sloan Digital Sky Survey Data Release 7 \citep[SDSS DR7;][]{Abazajian09} is 
a wide-field multiband imaging and spectroscopic survey which maps approximately 
one-quarter of the northern sky using a drift scanning technique.  Photometric 
data in the five-band SDSS system --- $u$, $g$, $r$, $i$, and $z$ --- is taken 
on a dedicated 2.5m telescope at the Apache Point Observatory in New Mexico 
\citep{Fukugita96,Gunn98}.  Galaxies with Petrosian $r$-band magnitudes 
$m_r < 17.77$ are selected for follow-up spectroscopic analysis \citep{Lupton01,
Strauss02}.  The galaxy colors are taken from the Korean Institute for Advanced 
Study Value-Added Galaxy Catalog \citep[KIAS-VAGC;][]{Choi10}, which contains 
galaxies from the SDSS DR7 main sky survey based on the New York University 
Value-Added Galaxy Catalog Data Release 7 \citep[NYU-VAGC;][]{Blanton05}.  Total 
star formation rates (SFR) and total specific star formation rates (sSFR) are 
from the MPA-JHU value-added catalog\footnote{Available at 
\url{http://www.mpa-garching.mpg.de/SDSS/DR7/}}, calculated using the technique 
described in \cite{Brinchmann04}.  The gas-phase chemical abundances used are 
from Douglass \& Vogeley (2017, in prep.), which are calculated using the Direct 
$T_e$ method.  The galaxies' large-scale environments are based on the void 
catalog compiled by \cite{Pan12}, which is constructed from SDSS DR7 using the 
VoidFinder algorithm of \cite{Hoyle02}.  Galaxies which are located within 
identified voids are labeled as void galaxies, and those which do not fall 
within a void are designated as a wall galaxy.  The large-scale environment 
within 5 \hMpc of the edge of the main SDSS DR7 footprint cannot be described 
using the VoidFinder algorithm because the minimum diameter of a void is 10 
\hMpc.  Therefore, the large-scale environment for galaxies within 5 \hMpc of 
the edge of the main footprint in SDSS DR7 is designated as unknown.

@@

Two galaxy samples are used in this analysis.  The largest one contains the 
11,845 dwarf galaxies spectroscopically observed in SDSS DR7.  The second sample 
contains those star-forming dwarf galaxies from the first set which have 
gas-phase chemical abundances in Douglass \& Vogeley (2017, in prep.), totaling 
2506 dwarf galaxies.  While this subset of galaxies contains only star-forming 
galaxies and is not representative of all dwarf galaxies, we show in Section 
\ref{sec:selection_effects} that this does not affect our results.

@@

In addition to comparing the target galaxies to their nearest neighbor, we also 
look at the relationship between the dwarf galaxies' properties and their 
proximity to the nearest galaxy group.  We use the Mr18 Berlind group catalog 
\cite{Berlind06} as our source of groups, which is built on the galaxies in the 
NYU-VAGC. The catalog identifies groups using a friends-of-friends algorithm 
\citep{Huchra82} to connect galaxies with $M_r < -18$.

@@

\subsection{Selection effects}\label{sec:selection_effects}

@@

% Absolute magnitude distribution of nearest neighbors
\begin{figure}
    \includegraphics[width=\textwidth]{Images/smallScaleEnvironment/Mr_distribution}
    \caption[Distribution of absolute magnitudes of nearest neighbors]{The 
    distributions in absolute magnitude (left panel) and absolute magnitude 
    relative to the target galaxy (right panel) of the nearest neighbor 
    galaxies.  The top row includes those nearest neighbors in units of \hMpc, 
    and the bottom row consists of the nearest neighbors in units of the 
    neighbor's virial radius.  Comparing the two plots in the right panel, we 
    see that the closest galaxy to the target dwarf galaxies is often of equal 
    or fainter magnitude than the target galaxy.  Alternatively, using the 
    virial radius of the neighbor galaxy as a measure of distance often finds a 
    brighter galaxy than the target galaxy, as the bottom right plot shows.}
    \label{fig:Mr_dist}
\end{figure}

For about 41\% of our dwarf galaxy sample, the two different metrics by which to 
define the nearest neighbor (minimum in units of \hMpc or virial radius of the 
nearest neighbor) return different neighboring galaxies.  Fig. \ref{fig:Mr_dist} 
compares the absolute magnitude distributions of these two nearest neighbor 
populations.  The left panel shows the distribution of the absolute magnitudes 
of the nearest neighbor galaxies, while the right panel shows the distribution 
in absolute magnitude of the nearest neighbor galaxy relative to its target 
galaxy.  The top row includes those nearest neighbors in units of \hMpc, and the 
bottom row consists of the nearest neighbors in units of the neighbor's virial 
radius.  Comparing the two plots in the right panel, we see that the closest 
galaxy to the target dwarf galaxy is often of equal or fainter magnitude than 
the target galaxy.  Alternatively, using the virial radius of the neighbor 
galaxy as a measure of the distance often finds a brighter galaxy than the 
target galaxy, as the bottom right plot in Fig. \ref{fig:Mr_dist} shows.  By 
using these two different distance metrics, we are able to probe the 
relationship between a dwarf galaxy and its nearest neighbor and nearest dark 
matter halo.

@@

\section[Conclusions]{Conclusions}

@@

Further analysis of this study should include comparing the properties of the 
target galaxies with the nearest neighbors' properties as a function of 
distance (``galactic conformity'').  The gas-phase chemical abundances between a 
galaxy and its nearest neighbor could be strongly correlated.
